JEThread: A Thread Pool Implementation in C

### 2018-05-04 thread-pool-implementation/paper.tex
on evaluation

\subsection{Work Stealing}
One of our goals is to achieve automatic load balancing. For this purpose, we implement work stealing for each worker thread. For each worker, we first do a sem\_trywait() to see if there is any job on its private queue. If there is none, then it randomly selects a victim thread and tries to steal a task from the victim and executes it. If keeps doing so until there is no more work to steal or if it has stolen $N$ times, where $N$ is the number of workers. This is our first attempt at work stealing. To optimize this, we could have kept randomly selecting victims until there is a victim with work above a certain threshold or up to a certain number of tries. To avoid all workers trying to steal when there is no tasks available, we can also use randomization to determine if a worker should try work stealing or just block until it is assigned a job. Another optimization we can perform is to steal more than one work at a time since we want it to be worth the overhead of running the algorithm. Although not reflected in the evaluation, we did try to evaluate the impact of the current work stealing techniques and found that even the naive approach improves JEThread's performance slightly.

\subsection{Internal Synchronization}
Each thread owns a privat task queue to store all its assigned tasks. Each queue is protected by either a mutex or a spinlock depending on users' specification. However, it is possible that a thread's task queue is empty when it is first initialized. It will waste CPU resource if such task is kept up and running without executing any tasks. Thus, we use semaphore to improve the efficiency of the thread pool. Thus we let the main thread invoking sema\_post on each thread once it finishes adding tasks to the thread's task queue to indicate the availability of tasks. However, the thread's task queue can be empty for other reasons. The most important one is that the thread finishes executing all assigned tasks. Thus, we implement work stealing algorithm to keep the idle thread busy by letting it trying to steal tasks from other threads by randomly choosing one. Since we don't want tasks stealing to create too much overhead as it tries to grab tasks from other threads, we simply let the idle thread using pthread\_mutex\_trylock() or pthread\_spin\_trylock(). If the stealing succeeds, the tasks is grabbed from the victim's task queue and executed. Otherwise, the steal will fail. If the steal fails too many times, that indicates many threads' task queue is already empty and the thread pool is getting close to running out of tasks. In this case, letting more threads trying to steal tasks will create unnecessary overhead, thus the thread will be blocked to prevent wasting resources. If eventually the threads run out of tasks, it will increase \textbf{wait\_sema} to let the main thread know it finishes its job. If all threads signify that they've finihsed their tasks, then thread pool is terminated.

@@

There are several goals we are aiming to achieve in the experiment. First, we want to run both long and short tasks in the threal pool to analyze which type of tests have the most impact on the performance on the program. We've created created various tests for the thread pool to perform and we will show the data collected later in the analysis section. We hope the low overhead of switch between tasks will help us outperform thread-based implementation as well go-based implementation. In addition, we implemented work stealing mechanism to improve the work throughput done by the thread pool. We hope the work stealing function will further improve the efficiency of the thread pool.

>>>>

\subsection{Work Stealing}
One of our goals is to achieve automatic load balancing. For this purpose, we implement work stealing for each worker thread. For each worker, we first do a sem\_trywait() to see if there is any job on its private queue. If there is none, then it randomly selects a victim thread. Since we don't want work stealing to create too much overhead as it tries to grab tasks from other threads, we first let the stealer try peeking at the victim's queue using pthread\_mutex\_trylock() or pthread\_spin\_trylock(). If the stealing succeeds, a task is grabbed from the victim's task queue and executed. Otherwise, the steal will fail. If the steal fails too many times, that indicates many threads' task queue is already empty and the thread pool is getting close to running out of tasks. In this case, letting more threads trying to steal tasks will create unnecessary overhead, thus the thread will be blocked to prevent wasting resources.

This is our first attempt at work stealing. To optimize this, we could have kept randomly selecting victims until there is a victim with work above a certain threshold or up to a certain number of tries. To avoid all workers trying to steal when there is no tasks available, we can also use randomization to determine if a worker should try work stealing or just block until it is assigned a job. Another optimization we can perform is to steal more than one work at a time since we want it to be worth the overhead of running the algorithm. Although not reflected in the evaluation, we did try to evaluate the impact of the current work stealing techniques and found that even the naive approach improves JEThread's performance slightly.

\subsection{Internal Synchronization}
Each worker thread owns a private task queue to store all its assigned tasks. Each queue is protected by either a mutex or a spinlock depending on users' specification. We highly recommend using a spinlock because we've ran experiments for both types of locks and observed that the spinlock always provides better performance. This is because the critical section is small and the code design rarely leads to lock contention.

In addition to preserving correctness through protected access to each worker's queue, we also try to keep JEThread's overhead low through efficient implementation. For example, when a worker runs out of tasks to execute, it can either busy wait or block. Because busy waiting will exhaust CPU resource, we decide to use a semaphore \textbf{task\_sema} to block a worker while there is no available task. When the main thread assigns tasks to the worker thread, it calls sema\_post on the worker's task\_sema.

In order to wait for all worker threads to finish, in thread\_pool\_wait() we first mark each worker's private state \textbf{count} to true, then wake up each worker using \textbf{task\_sema} , and wait on each worker's \textbf{wait\_sema}. Once a worker wakes up, if there is no more job on its private queue and \textbf{count} is true, it wakes up the main thread to indicate that it has finished the current set of tasks.

@@

There are several goals we are aiming to achieve in the experiment. Firstly, we want to test the flexibility of the thread pool by running a mixture of long(blocking) and short tasks and evaluate how we perform compared to other implementations. Secondly, we want tot
